Sparse Representation Approaches for First Stage Retrieval: A Review

### 2022-01-06 src/article.tex
Use dot at the end of the paragraph name.
Add training paragraph for SNRM.

    \paragraph{Loss function.}
    The loss function for the $i^{th}$ training instance has the following form:
    \begin{equation}
        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi(q_i)||\psi(d_{i1})||\psi(d_{i2}))
    \end{equation}
    where $\mathcal{L}$ - hinge loss \cite{IRNLP}, $q_i$ - query, $d_{i1}$ and $d_{i2}$ - two
    document candidates, $y_i \in \{-1, 1\}$ - label that shows the relevance between the document
    and the query, $\lambda$ - the hyper-parameter to control the sparsity of the learned
    representations, $L_1$ is norm (i.e.
    \begin{math}
        L_1(\vec{\upsilon}) = \sum_{i=1}^{|\vec{\upsilon}|}|\vec{\upsilon}_i|
    \end{math}
    ) and $||$ is concatenation sign.\\\par

@@

    \paragraph{Architecture}
    \begin{figure}[h]
        \centering
        \def\svgwidth{\columnwidth}
        \input{SPLADE.pdf_tex}
        \caption{Final representation calculation with log saturation}
        \label{fig:SPLADE}
    \end{figure}
    The importance estimation formula Equation (\ref{eqn:sparterm:importance-predictor})
    has been modified by adding a log-saturation effect (Figure \ref{fig:SPLADE}), in order to
    prevent the dominance of some terms to dominate and naturally ensures sparsity in
    representations:
    \begin{equation}
        \label{eqn:splade:importance}
        w_j=\sum_{i \in t}log(1+ReLU(w_{ij}))
    \end{equation}

    \paragraph{Loss function}
    The function presented in the Equation (\ref{eqn:sparterm:loss}) is supplemented with negative
    sampling as shown below:
    \begin{equation}
        L_{rank-IBN}=-log
        \frac{
            e^{s(q_i,d_i^+)}
        }{
            e^{s(q_i,d_i^+)} +
            e^{s(q_i,d_i^-)} +
            \sum_j e^{s(q_i,d_{i,j}^-)}
        }
    \end{equation}
    where IBN is in-batch negatives sampling strategy, which is effective for learning first-stage
    rankers \cite{IBN1,IBN2,IBN3}, $q_i$ is query, $d_i^+$ and $d_i^-$ are positive and negative
    documents respectively, ${d_{i,j}^-}$ is set of positive documents from other queries,
    but negative for the current one \cite{SPLADE}.

@@

    \paragraph{Architecture}
    \begin{figure}[h]
        \centering
        \includegraphics[scale=0.2]{RepBERT.png}
        \caption{RepBERT architecture}
        \label{fig:RepBERT}
    \end{figure}
    The process of receiving embeddings from document or query sentences is shown
    in the Figure \ref{fig:RepBERT}.

>>>>

    \paragraph{Loss function.}
    The loss function for the $i^{th}$ training instance has the following form:
    \begin{equation}
        \label{eqn:snrm:loss-function}
        \mathcal{L}(q_i,d_{i1},d_{i2},y_i) + \lambda\;L_1(\phi(q_i)||\psi(d_{i1})||\psi(d_{i2}))
    \end{equation}
    where $\mathcal{L}$ - hinge loss \cite{IRNLP}, $q_i$ - query, $d_{i1}$ and $d_{i2}$ - two
    document candidates, $y_i \in \{-1, 1\}$ - label that shows the relevance between the document
    and the query, $\lambda$ - the hyper-parameter to control the sparsity of the learned
    representations, $L_1$ is norm (i.e.
    \begin{math}
        L_1(\vec{\upsilon}) = \sum_{i=1}^{|\vec{\upsilon}|}|\vec{\upsilon}_i|
    \end{math}
    ) and $||$ is concatenation sign.

    \paragraph{Training.}
    Weak supervision is used to train the model.

    Adam is selected as the optimizer.
    The learning rate is chosen from
    ${\{5 \cdot 10^{-5}, 1 \cdot 10^{-4}, 5 \cdot 10^{-4}, 1 \cdot 10^{-3}, 5 \cdot 10^{-3}\}}$,
    the batch size from \{32, 64, 128\}.
    Ngram representation network has n equal to 5 and two or three hidden layers.
    Hidden layer sizes and the output layer sizes are selected from \{100, 300, 500\} and
    \{5k, 10k, 20k\}, respectively.
    The hyper-parameter $\lambda$ (see Equation (\ref{eqn:snrm:loss-function})) is chosen from the
    $[5 \cdot 10^{-9}, 1 \cdot 10^{-7}]$ interval.
    The dropout rate is selected from \{0.6, 0.8, 1\}.\\\par

@@

    \paragraph{Architecture.}
    \begin{figure}[h]
        \centering
        \def\svgwidth{\columnwidth}
        \input{SPLADE.pdf_tex}
        \caption{Final representation calculation with log saturation}
        \label{fig:SPLADE}
    \end{figure}
    The importance estimation formula Equation (\ref{eqn:sparterm:importance-predictor})
    has been modified by adding a log-saturation effect (Figure \ref{fig:SPLADE}), in order to
    prevent the dominance of some terms to dominate and naturally ensures sparsity in
    representations:
    \begin{equation}
        \label{eqn:splade:importance}
        w_j=\sum_{i \in t}log(1+ReLU(w_{ij}))
    \end{equation}

    \paragraph{Loss function.}
    The function presented in the Equation (\ref{eqn:sparterm:loss}) is supplemented with negative
    sampling as shown below:
    \begin{equation}
        L_{rank-IBN}=-log
        \frac{
            e^{s(q_i,d_i^+)}
        }{
            e^{s(q_i,d_i^+)} +
            e^{s(q_i,d_i^-)} +
            \sum_j e^{s(q_i,d_{i,j}^-)}
        }
    \end{equation}
    where IBN is in-batch negatives sampling strategy, which is effective for learning first-stage
    rankers \cite{IBN1,IBN2,IBN3}, $q_i$ is query, $d_i^+$ and $d_i^-$ are positive and negative
    documents respectively, ${d_{i,j}^-}$ is set of positive documents from other queries,
    but negative for the current one \cite{SPLADE}.

@@

    \paragraph{Architecture.}
    \begin{figure}[h]
        \centering
        \includegraphics[scale=0.2]{RepBERT.png}
        \caption{RepBERT architecture}
        \label{fig:RepBERT}
    \end{figure}
    The process of receiving embeddings from document or query sentences is shown
    in the Figure \ref{fig:RepBERT}.
